---
title: "Hayden and Ying R and ggplot"
author: "Hayden Wainwright"
date: "3/23/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

## Initial tutorial in R and ggplot

### Hayden and Ying

### Overview:

Research Question:

Does frog mortality rate change in current COVID-19 situation in Ontario?

H0: nothing will change
H1: Frog mortality rate decreases with decreasing traffic under Ontario lockdown 
H2: Frog mortality rate increase with increasing human predation for food under pandemic


### Topics to be dicussed

#### R Basics (5-10min tops) (Hayden):
1. how to make a project
2. markdown vs script
3. github links
4. handling error messages 
        (dependancies missing, unexpect ",", googling answers)
5. how to google 101
6. other helpful links (Ying)
7. packages

#### Simulating work (30min) (Hayden):
1. how to generate a "fake" data set
  (date, frog species, frogs dead on road, estimated pop size, location, sampling year, Avg or total cars on road per unit time, food supply in Ontario as a proxy for human needs for food)
2. how to export the data set
3. power analysis????? (can go if need be)




#### What is power analysis? (Ying)

```{r, echo=FALSE}
# How to make a nice table? Please check: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html 

library(knitr)
library(kableExtra)

PAcol1 <- c("", "Null hypothesis H0 is true", "Alternative hypothesis H1 is ture")
PAcol2 <- c("Fail to reject H0", "Good", "Type II error")
PAcol3 <- c("Reject H0", "Type I error", "Power")
PAtable <- data.frame(PAcol1, PAcol2, PAcol3)
colnames(PAtable) <- NULL

kable(PAtable) %>%
  kable_styling(c("striped", "hover"), full_width = FALSE, position = "float_right")
```

**Statistical power** is the probability that the test correctly rejects the null hypothesis.

**Power** = Pr (Reject H0 | H1 is true) = 1 - Type II error rate

It mostly depends on:

* **Alpha level** - significance level = Type I error rate. Larger alpha, greater power.
    + Pr (observed | H0 is ture) < alpha, reject null 
    + Pr (observed | H0 is ture) > alpha, fail to reject null 

* **Effect size** - quantified magnitude of a result present in the population. Larger effects, larger power. 

* **Sample size** - Larger sample size, smaller sampling error, greater statistical power. 
    + *Paper recommended:  Sullivan and Feinn. 2002. Using Effect Size—or Why the P Value Is Not Enough*

**A power analysis** can be used to estimate the minimum sample size required for an experiment, given a desired significance level, effect size, and statistical power.


##### How to do power analysis? (Ying)






#### dplyr (20 min) (Ying):
1. Import Data
2. How to change data into what you want
3. how to get rid of NA in your data
4. how to make a new collumn in your data
5. how to change the orientation and layout of your data (Hayden)
6. double checking data sets for spelling mistakes
7. piping and filtering/sort data
8. creating new data frames
9. how to make a small program (for loop, if statements, make your own functions) (Hayden to help)



#### ggplot (20min) (Ying and Hayden):
1. qplot 
2. plotting basic functions in ggplot using geoms
    + boxplot plot: mortality vs year 
3. adding trend lines 
    + scatter plot + linear model: mortality vs traffic 
4. adding multiple data sets to same graph
5. anotate text and axises
6. changin colour schemes and shapes
    + fill color of different frog species, different locations
7. change theme of a plot 
8. plotting models 
9. confidence intervals and standard error bars
    + linear model: mortality vs traffic, mortality vs food supply in ontario 
10. how to add multiple plots to one graph space





